Inference for multivariate peaks over threshold models

### 2021-10-27 methodology_projgamma.tex
overleaf updates

\subsection{Projected gamma family}
  A natural  distribution to consider in ${\mathbb R}^p_+$ is given by a product of independent
  univariate Gamma distributions. Let
    $\bm{ X} \sim \prod_{\ell = 1}^d\text{Ga}\left(X_{\ell}\mid\alpha_{\ell},\beta_{\ell}\right)$, 
    $a_\ell$ and $b_\ell$ are the shape and scale parameters, repectively. Using Equations (\ref{eqn:pnormt}) -- (\ref{eqn:pnormjac}) we have the joint density
  \begin{equation*}
    f(r,\bm{ y}) = \prod_{\ell = 1}^{d}
      \left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}(ry_{\ell})^{\alpha_{\ell} - 1}
          \exp\lbrace-\beta_{\ell}ry_{\ell}\rbrace\right]
      \times r^{d-1}\left[y_d -
            {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right],
  \end{equation*}
  Integrating out $r$ yields the resulting \emph{Projected Gamma} density
  \begin{equation}
    \label{eqn:projgamma}
    \text{PG}(\bm{ y}\mid\bm{ \alpha},\bm{ \beta}) =
          \prod_{\ell = 1}^d\left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}
                y_{\ell}^{\alpha_{\ell} - 1}\right]
      \times \left[y_d -
          {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right]
      \times \frac{\Gamma({\textstyle\sum}_{\ell = 1}^d\alpha_{\ell})}{\left({\textstyle\sum}_{\ell = 1}^d
                    \beta_{\ell}y_{\ell}\right)^{{\scriptstyle\sum_{\ell = 1}^d \alpha_{\ell}}}}
  \end{equation}
  defined for $\bm{y}\in \mathcal{S}_p^{d-1}$, and for any $p>0$. To avoid idetifiability problems when estimating the shape and scale parameters, we set $\beta_1 = 1$.
  \cite{nunez2019} obtain the density in Equation (\ref{eqn:projgamma})
  for $p=2$ as a multivariate distribution for directional data, using spherical coordinates.
  For $\bm{ y}\in \mathcal{S}_1^{d-1}$, and $\beta_{\ell} = \beta, \forall\ell$, the density in Equation (\ref{eqn:projgamma}) corresponds to that of a Dirichlet distribution.

  The projected gamma family is simple to specify and has very tractable computational properties. Thus, we use it as a building block for the angular measure $\Phi$ models. To build a flexible family of distributions in $\mathcal{S}_p^{d-1}$ we consider mixtures of projected gamma densities defined as
  \begin{equation} \label{PGmix}     
     f(\bm{y}) = \int_\Theta \text{PG}(\bm{y}|\bm{\theta}) dG(\bm{\theta}), 
  \end{equation} 
  where $\bm{\theta} = (\bm{\alpha}, \bm{\beta})$. Following a Bayesian non-parametric approach \citep{Ferguson74,Antoniak1974} we assume that $G$ is drawn from a random measure. In particular, assuming a Dirichlet process prior, we have a hierarchical formulation of the mixture model given by
  \[    \bm{y} \sim \text{PG}(\bm{y}|\bm{\theta}) , \;\;\; \bm{\theta} \sim G, \;\;\; G \sim \text{DP}(\eta, G_0),\]
  where DP denotes a Dirichlet process, $\eta$ is the precision parameter, and $G_0$ is the centering distribution.

  Our strategy consists of describing the angular distribution $\Phi$ using a sample based approach with the following steps: (i) Apply the transformation in Equation (\ref{eqn:standardization}) to the original data; (ii) Obtain the subsample of the standardized observations that satisfy $R>1$; (iii) Take a finite $p$ and project the observations onto $\mathcal{S}_p^{d-1}$; (iv) Fit the model in Equation (\ref{PGmix}) to the resulting data and obtain samples from the fitted model; (v) Project the resulting samples onto $\mathcal{S}_\infty^{d-1}$.
  For step (iv) we use a Bayesian approach that is implemented using a purposely developed Markov chain Monte Carlo that is described in the next section. 
  {\bf Instead of the olf Figure 1, it would be nice to add a figure where you illustrate this procedure. Give it a try.}
%  The full conditional for $r$ takes the form
%  \begin{equation}
%    \label{eqn:rfullcond}
%    r\mid\bm{ \alpha},\bm{ \beta}, y \sim \text{Ga}\left(r\mid{\textstyle\sum}_{\ell = 1}^d
%                    \alpha_{\ell}, {\textstyle\sum}_{\ell = 1}^d \beta_{\ell}y_{\ell}\right).
%  \end{equation}

\subsubsection{Inference for the projected gamma mixture model}
{\bf Our focus is on model (\ref{PGmix}). Describe how you build a sample based approach to fitting that model. How you introduce latent variables, both to break the mixture and to produce the PG. You can then discuss variations that essentially consists on different priors. So, basically, you need to condense the next three pages in about a page.}

We consider a basic model that assumes all observations are generated by a single projected Gamma
  distribution.  The shape and rate parameters for the latent \text{Gamma} random variables,
  $\alpha_{\ell},\beta_{\ell}$, must be greater than 0, so a logical prior to use would be a Gamma prior for
  each parameter, excepting for $\beta_1 := 1$.
  \begin{equation}
    \label{eqn:vanillap}
    \begin{aligned}
      \bm{ y}\mid\alpha,\beta &\sim \text{PG}(\bm{ y}\mid\alpha,\beta)\\
      \bm{ \alpha},\bm{\beta} &\sim {\textstyle \prod}_{\ell = 1}^d \text{Ga}(\alpha_{\ell} \mid \xi_{\ell},\tau_{\ell})
              \times {\textstyle \prod}_{\ell = 2}^d \text{Ga}(\beta_{\ell}\mid \zeta_{\ell},\sigma_{\ell}).
    \end{aligned}
  \end{equation}
  A useful property of the Projected Gamma family is the latent independent Gammas assumption.  Recall
  in Eqn.~\ref{eqn:protopg}, we have a density defined as a product of independent Gamma densities,
  multiplied by a factor of $r^{d-1}C$, where $C$ is constant with respect to the parameters
  $\alpha,\beta$, as well as $r$.  If we conduct inference via a Gibbs sampler, then by simulating
  this latent $r$ we are able to perform inference of $(\alpha_{\ell},\beta_{\ell})$ independent of
  that of $(\alpha_k,\beta_k)$.  The full conditional for $r$ is described in Eqn.~\ref{eqn:rfullcond}.
  The full conditional for $\beta_{\ell}$, $l = 2,\ldots,d$ then becomes
  \begin{equation}
    \label{eqn:betafc}
    \beta_{\ell}\mid \bm{ y}, r, \alpha_{\ell} \sim \text{Ga}\left(n + \alpha_{\ell} + \zeta_{\ell},
                                      {\textstyle \sum}_{i = 1}^nr_iy_{il} + \sigma_{\ell}\right).
  \end{equation}
  That is, given $\alpha_{\ell}$, $r$, we can sample $\beta_{\ell}$ directly as a Gamma random
  variable.  In the sampling for $\alpha_{\ell}$, we can integrate $\beta_{\ell}$ out leaving us a
  quantity proportional to a density, which we can sample via the Metropolis Hastings
  algorithm\citep{hastings1970}.  For $\alpha_{\ell}$, $\ell = 2,\ldots, d$, This takes the form
  \begin{equation}
    \label{eqn:alphafc}
    f(\alpha_{\ell} \mid \bm{ y}, r) \propto
    \frac{\left({\textstyle \prod}_{i = 1}^nr_iy_{il}\right)^{\alpha_{\ell} - 1}}{\Gamma^n(\alpha_{\ell})} \times
    \alpha_{\ell}^{\xi_{\ell} - 1}\exp\{-\tau_{\ell}\alpha_{\ell}\} \times
    \frac{\Gamma(n\alpha_{\ell} + \zeta_{\ell})}{
        \left({\textstyle\sum}_{i = 1}^n r_iy_{il} + \sigma_{\ell}\right)^{(n * \alpha_{\ell} + \zeta_{\ell})}}.
  \end{equation}
  For $\alpha_1$, as we do not integrate out $\beta_1$ as it is defined to be $1$, this takes the
  slightly simpler form
  \begin{equation}
    \label{eqn:alphafc1}
    f(\alpha_1 \mid \bm{ y}, r) \propto
      \frac{{\textstyle\prod}_{i = 1}^n ry_{i1}^{\alpha_1 - 1}}{\Gamma^n(\alpha_1)} \times
      \alpha_1^{\xi_1 - 1}\exp\{-\tau_1\alpha_1\}.
  \end{equation}
  The \emph{projected restricted Gamma} model sets $\beta_{\ell} := 1 \forall \ell \in \lbrace 1,\ldots,d\rbrace$,
  so in that model inference on all $\alpha_{\ell}$ will use the simpler form.

\subsubsection{Finite Mixture Model}
The projected Gamma distribution offers a flexible model for describing direction.  However, it is
  conceivable that a single Projected Gamma distribution will not sufficiently capture the
  variability of real data.  In that case, a logical extension to make to the above model is to
  assume the data, $\bm{ y}$, are generated from a mixture of projected Gammas.  We consider a finite
  mixture,
  \begin{equation}
    \label{eqn:fmixp}
    \begin{aligned}
      \bm{ y}_i &\sim \sum_{j = 1}^J\pi_j\text{PG}\left(\bm{ y}\mid \bm{ \alpha}_j, \bm{ \beta}_j\right)\\
      \bm{ \alpha}_j &\sim {\textstyle \prod}_{\ell = 1}^d \text{Ga}\left(\alpha_{\ell}\mid\xi_{\ell},\tau_{\ell}\right)\\
      \bm{ \beta}_j &\sim {\textstyle \prod}_{\ell = 2}^d \text{Ga}\left(\beta_{\ell}\mid\zeta_{\ell},\sigma_{\ell}\right)
    \end{aligned}
    \hspace{2cm}
    \begin{aligned}
      \bm{ \xi},\bm{\tau} &\sim {\textstyle \prod}_{\ell = 1}^d \text{Ga}(\xi_{\ell}\mid a,b) \times \text{Ga}(\tau_{\ell}\mid c,d)\\
      \bm{ \zeta},\bm{\sigma} &\sim {\textstyle\prod}_{\ell = 2}^d\text{Ga}(\zeta_{\ell} \mid a,b)\times \text{Ga}(\sigma_{\ell}\mid c,d)\\
      \bm{ \pi} &\sim \text{Dir}(\pi_0)
    \end{aligned}
  \end{equation}
  In addition to allowing multiple mixture components, we assume those mixture components themselves
  descend from a product of independent Gamma densities. The standard method for sampling from a
  mixture distribution includes a data augmentation step, generating a
  $\delta_i \in \lbrace 1,\ldots, J\rbrace$ indicating to which mixture component a
  particular observation comes from, with
  \begin{equation*}
    \text{P}\left[\delta_i = j\mid\alpha_j,\beta_j\right] \propto \pi_j\text{PG}\left(\bm{ y}_i\mid \bm{ \alpha}_j,\bm{ \beta}_j\right).
  \end{equation*}
  By sampling and conditioning on $\delta_i$, we end up with very similar semi-conjugate forms for
  inference of $\bm{ \alpha}_j$, $\bm{ \beta}_j$ that were used in the basic model.  The changes made
  to Equations~\ref{eqn:betafc},\ref{eqn:alphafc},\ref{eqn:alphafc1} are such that we're conducting inference
  on a mixture component given the mixture component identifiers $\bm{ \delta}$, so instead of indexing
  $i$ over the whole data set, we're indexing $i$ over the set $\{i :\delta_i = j\}$, with $n = n_j$,
  the cardinality of that set. Inference on $\bm{ \xi,\tau,\zeta,\sigma}$ follows a similar form as
  that of $\bm{ \alpha,\beta}$---they are Gamma-Gamma relationships, so rate parameters have a conjugate
  Gamma posterior, and we can integrate them out to perform inference on the shape parameters via
  Metropolis Hastings.

Additionally, we investigate a finite mixture extension of the \emph{Projected Restricted Gamma} model,
  where we define all $\beta_{\ell} := 1$, for $l = 1,\ldots,d$, for $j = 1,\ldots,J$.  This affects the above
  model by removing inference steps for $\bm{ \beta},\bm{ \zeta},\bm{ \sigma}$.  The inference for
  $\alpha_{\ell}$ for all $\ell = 1,\ldots,d$ follows the form in Equation~\ref{eqn:alphafc1}; inference on
  $\bm{ \xi,\tau}$ follows the same form as the unrestricted model.  The impetus for the restricted
  gamma model is that in the context of a mixture model, the additional flexibility offered by the
  unrestricted Gamma model may be better represented with additional clusters.

\subsubsection{Dirichlet Process Mixture Model}
A logical extension to the finite mixture model described in Equation~\ref{eqn:fmixp} would be to assume the
  $\bm{ y}$'s descending from an infinite mixture model.  This to some extent solves the issue of
  selecting the number of mixture components, as that will effectively be done so in the model.
  \begin{equation}
    \label{eqn:infmixp}
    \begin{aligned}
      \bm{ y}_i &\sim \sum_{j = 1}^J\pi_j\text{PG}\left(\bm{ y}\mid \bm{ \alpha}_i, \bm{ \beta}_i\right)\\
      (\alpha_i,\beta_i) &\sim \text{DP}\left(\eta, G_0\right)\\
      &~\hspace{-2cm}G_0 = {\textstyle\prod}_{\ell = 1}^d\text{Ga}\left(\alpha_{jl}\mid\xi_{l},\tau_{l}\right)
                    \times{\textstyle\prod}_{\ell = 2}^d\text{Ga}\left(\beta_{jl}\mid\zeta_{l}\tau_{l}\right)
    \end{aligned}
    \hspace{1cm}
    \begin{aligned}
      \bm{ \xi},\bm{ \tau} &\sim {\textstyle \prod}_{\ell = 1}^d \text{Ga}(\xi_{\ell}\mid a,b)\times \text{Ga}(\tau_{\ell}\mid c,d)\\
      \bm{ \zeta},\bm{\sigma} &\sim {\textstyle\prod}_{\ell = 2}^d\text{Ga}(\zeta_{\ell} \mid a,b)\times \text{Ga}(\sigma_{\ell}\mid c,d)\\
      \bm{ \eta} &\sim \text{Ga}(\eta \mid 2, 0.1).
    \end{aligned}
  \end{equation}
  To fit the Dirichlet process, we use a variant of the collapsed Gibbs sampler detailed as
  Algorithm~8 in \citet{neal2000}.  In this algorithm, we perform a data augmentation to create a
  cluster identifier, $\delta_i$, for each observation. Let $J$ indicate the number of extant clusters.
  Under the collapsed sampler scheme, for existing clusters, the probability of cluster membership is
  calculated as
  \begin{equation}
    \label{eqn:dpdelta}
    \text{P}\left(\delta_i = j\mid r, \bm{ y_i} \bm{ \alpha}, \bm{ \beta}\right) \propto
        \frac{n_j^{\not i}}{\sum_{k = 1}^{J} n_k + \eta}
        {\textstyle\prod}_{\ell = 1}^d\text{Ga}\left(r_iy_{il}\mid\alpha_{jl},\beta_{jl}\right).
  \end{equation}
  where $n_j^{\not i} = {\textstyle\sum}_{k\neq i}1_{k == j}$.  Effectively, if the observation under review is a
  singleton, its existing cluster is destroyed.  For new clusters, ostensibly we would integrate out
  cluster parameters to achieve a posterior predictive density; though doing so in a
  non-conjugate model may be difficult.  Instead, the impetus for using this specific DP fitting algorithm
  is it simulates the posterior predictive density by generating $m$ new candidate clusters from
  their respective priors given $\bm{ \xi,\tau,\zeta,\sigma}$.  Then the probability of $y_i$
  belonging to a particular candidate cluster is calculated as
  \begin{equation}
    \label{eqn:dpdeltanew}
    \text{P}\left[\delta_i = k\mid r, \bm{ \alpha}_k^{\prime}, \bm{ \beta}_k^{\prime}\right] \propto
        \frac{\eta / m}{\sum_{j = 1}^{J} n_j + \eta}
        {\textstyle\prod}_{\ell = 1}^d\text{Ga}\left(r_iy_{il}\mid\alpha_{kl}^{\prime},\beta_{kl}^{\prime}\right),
  \end{equation}
  where $(\cdot)^{\prime}$ indicates a value as belonging to a new candidate cluster, and $k$ indexes
  along candidate clusters.  If a new cluster is selected, then
  $(\bm{\alpha}_{J+1},\bm{\beta}_{J+1}) = (\bm{\alpha}_k^{\prime},\bm{\beta}_k^{\prime})$, and $J = J + 1$.

Inference upon $\bm{ \alpha}_j,\bm{ \beta}_j$ operates in the same manner as with the finite mixture
  model.  For $\bm{ \xi,\tau,\zeta,\sigma}$, we have $J$ cluster components, indicating $J$
  observations that again are under a gamma-gamma model.  The rate parameters have a previously
  described conjugate posterior, that we can integrate out to more directly sample the shape
  parameters.  As before, we will be looking at this model, and the restricted gamma form of this model, where
  $\beta_{\ell} := 1$ for $\ell = 1,\ldots,d$.  This removes inference steps for $\bm{ \zeta},\bm{ \sigma}$
  as well.

\subsubsection{A Log-Normal Prior for Shape Parameters}
A final extension we can make to this model is to allow for some structure, or dependence, in the prior
  for shape parameters within a cluster.  The impetus for this approach is from recognizing that the
  algorithm we are using for sampling the DP may not adequately represent the posterior density for
  a new cluster, as it is assuming sampling from an independent prior.  If we account for dependence
  between dimensions within the prior, then we may be able to more efficiently sample new clusters.
  As shape parameters must be greater than 0, it is logical to assume a log normal prior, as that
  allows us the familiar forms of a mean vector and covariance matrix for representing the source of
  cluster parameters. So,
  \begin{equation}
    \label{eqn:dpln}
    \begin{aligned}
      \bm{ y}_i &\sim \sum_{j = 1}^J\pi_j\text{PG}\left(\bm{ y}\mid \bm{ \alpha}_i, \bm{\beta}_i\right)\\
      (\bm{\alpha}_i, \bm{\beta}_i) &\sim \text{DP}\left(\eta, G_0\right)\\
        &~\hspace{-1cm}G_0 = \mathcal{N}\left(\log\bm{ \alpha}_{j}\mid\mu,\Sigma\right)\times
            {\textstyle\prod}_{\ell = 2}^d\text{Ga}\left(\beta_{j,\ell}\mid\zeta_{\ell},\sigma_{\ell}\right)
    \end{aligned}
    \hspace{1cm}
    \begin{aligned}
      \mu,\Sigma &\sim \mathcal{N}\left(\mu\mid\mu_0,\Sigma_0\right) \times \text{IG}\left(\nu,\Psi\right)\\
      \bm{ \zeta},\bm{\sigma} &\sim {\textstyle\prod}_{\ell = 2}^d\text{Ga}(\zeta_{\ell} \mid a,b) \times \text{Ga}(\sigma_{\ell} \mid c,d) \\
      \bm{ \eta} &\sim \text{Ga}(\eta \mid 2, 0.1).
    \end{aligned}
  \end{equation}
  We hope that this will allow us to better incorporate the relationship between dimensions in
  sampling new clusters, allowing for better mixing of the model.  The downside is one of computational
  complexity---this introduces a $d$-dimensional normal distribution into the model, requiring a
  $d$-dimensional matrix inversion step.  As matrix inversion is computationally costly, this limits
  the potential number of dimensions possible.  The posterior inference strategies for $\beta,\zeta,\sigma$
  are the same as for prior models and have been described.  For inference on $\alpha_j$ given
  $r, \mu,\Sigma$, we rely on a $d$-dimensional Metropolis Hastings step, where the log-posterior
  density is proportional to
  \begin{equation}
    \label{eqn:dplnlp}
    \begin{aligned}
    f(\bm{\alpha}_j\mid\bm{\delta},\bm{y},\mu,\Sigma,\zeta,\sigma) &\propto
    \frac{\left(\prod_{i:\delta_i = j}y_{i,\ell}\right)^{\alpha_{j,\ell} - 1}}{\Gamma(\alpha_{j,\ell})^{n_j}}
        \prod_{\ell = 2}^d\left[\frac{\left(\prod_{i:\delta_i = j}y_{i,\ell}\right)^{\alpha_{j,\ell} - 1}
        \Gamma(n_j\alpha_{j,\ell} + \zeta)}{\Gamma(\alpha_{j,\ell})^{n_j}\left(\sum_{i:\delta_i = j}y_{i,\ell}
        + \sigma\right)^{n_j\alpha_{j,\ell} + \zeta}}\right]\\
        &\hspace{1cm}\times
        \lvert \Sigma\rvert^{-\frac{1}{2}}\prod_{\ell = 1}^d\alpha_{j,\ell}^{-1}\exp\left\lbrace
          -\frac{1}{2}(\log\alpha_j - \mu)^T\Sigma^{-1}(\log\alpha_j - \mu)\right\rbrace.

>>>>

\subsection{Projected gamma family\label{subsec:projgamma}}
  A natural  distribution to consider in ${\mathbb R}^p_+$ is given by a product of independent
  univariate Gamma distributions. Let
    $\bm{ X} \sim \prod_{\ell = 1}^d\text{Ga}\left(X_{\ell}\mid\alpha_{\ell},\beta_{\ell}\right)$, 
    $\alpha_\ell$ and $\beta_\ell$ are the shape and scale parameters, respectively. Using Equations (\ref{eqn:pnormt}) -- (\ref{eqn:pnormjac}) we have the joint density
  \begin{equation*}
    f(r,\bm{ y}) = \prod_{\ell = 1}^{d}
      \left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}(ry_{\ell})^{\alpha_{\ell} - 1}
          \exp\lbrace-\beta_{\ell}ry_{\ell}\rbrace\right]
      \times r^{d-1}\left[y_d +
            {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right],
  \end{equation*}
  Integrating out $r$ yields the resulting \emph{Projected Gamma} density
  \begin{equation}
    \label{eqn:projgamma}
    \text{PG}(\bm{ y}\mid\bm{ \alpha},\bm{ \beta}) =
          \prod_{\ell = 1}^d\left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}
                y_{\ell}^{\alpha_{\ell} - 1}\right]
      \times \left[y_d +
          {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right]
      \times \frac{\Gamma({\textstyle\sum}_{\ell = 1}^d\alpha_{\ell})}{\left({\textstyle\sum}_{\ell = 1}^d
                    \beta_{\ell}y_{\ell}\right)^{{\scriptstyle\sum_{\ell = 1}^d \alpha_{\ell}}}}
  \end{equation}
  defined for $\bm{y}\in {\mathbb S}_p^{d-1}$, and for any $p>0$. To avoid identifiability problems when estimating the shape and scale parameters, we set $\beta_1 = 1$.
  \cite{nunez2019} obtain the density in Equation (\ref{eqn:projgamma})
  for $p=2$ as a multivariate distribution for directional data, using spherical coordinates.
  For $\bm{ y}\in {\mathbb S}_1^{d-1}$, and $\beta_{\ell} = \beta$ for all $\ell$, the density in Equation (\ref{eqn:projgamma}) corresponds to that of a Dirichlet distribution.

  The projected gamma family is simple to specify and has very tractable computational properties. Thus, we use it as a building block for the angular measure $\Phi$ models. To build a flexible family of distributions in ${\mathbb S}_p^{d-1}$ we consider mixtures of projected gamma densities defined as
  \begin{equation} \label{eqn:PGmix}     
     f(\bm{y}) = \int_\Theta \mathcal{PG}(\bm{y}|\bm{\theta}) dG(\bm{\theta}), 
  \end{equation} 
  where $\bm{\theta} = (\bm{\alpha}, \bm{\beta})$. Following a Bayesian non-parametric approach \citep{Ferguson74,Antoniak1974} we assume that $G$ is drawn from a random measure. In particular, assuming a Dirichlet process prior, we have a hierarchical formulation of the mixture model that, for a vector of observations $\bm{y}_i$, is given by
  \begin{equation}\label{eqn:dppg}    \bm{y}_i \sim \text{PG}(\bm{y}_i|\bm{\theta}_i) , \;\;\; \bm{\theta}_i \sim G, \;\;\; G \sim \mathcal{DP}(\eta, G_0),
  \end{equation}
  where $\mathcal{DP}$ denotes a Dirichlet process, $\eta$ is the precision parameter, and $G_0$ is the centering distribution.

  Our strategy consists of describing the angular distribution $\Phi$ using a sample based approach with the following steps: (i) Apply the transformation in Equation (\ref{eqn:standardization}) to the original data; (ii) Obtain the subsample of the standardized observations that satisfy $R>1$; (iii) Take a finite $p$ and project the observations onto ${\mathbb S}_p^{d-1}$; (iv) Fit the model in Equation (\ref{eqn:PGmix}) to the resulting data and obtain samples from the fitted model; (v) Project the resulting samples onto ${\mathbb S}_\infty^{d-1}$.
  For step (iv) we use a Bayesian approach that is implemented using a purposely developed Markov chain Monte Carlo that is described in the next section. 
  {\bf Instead of the old Figure 1, it would be nice to add a figure where you illustrate this procedure. Give it a try.}

\subsubsection{Inference for the projected Gamma mixture model}
The Dirichlet process mixture model groups observations into stochastically assigned \emph{clusters}.  Building out the methods of inference for Equation~\ref{eqn:dppg}, let $n_j^{-i}$ be the number of observations in cluster $j$ not including observation $i$.  Let $J^{-i}$ be the number of extant observations, not including any singleton that observation $i$ is in. Under this model, the probability of cluster membership is proportional to
\begin{equation*}
    \text{P}\left[\delta_i = j\mid\ldots\right] \propto \begin{cases}
        n_j^{-i}\mathcal{PG}\left(\bm{y}_i\mid\bm{\alpha}_j,\bm{\beta}_j\right)\hspace{0.5cm} &\text{~for~}j \in \lbrace 1,\ldots,J^{-i}\rbrace\\
        \eta\int\mathcal{PG}\left(\bm{y}_i\mid\bm{\alpha}_j,\bm{\beta}_j\right)dG_0(\bm{\alpha}_j,\bm{\beta}_j)\hspace{0.5cm} &\text{~for~}j = J^{-i} + 1.
        \end{cases}
\end{equation*}
Of course, if $G_0$ is not a conjugate prior for the kernel distribution, then the integral is difficult.  Instead, using algorithm 8 from \cite{neal2000}, we draw $m$ candidate clusters, $\bm{\alpha}_j,\bm{\beta}_j$ for $j = J^{-i} + 1,\ldots, J^{-i} + m$ from $G_0$. Then, the probability of cluster membership is proportional to
\begin{equation}
    \text{P}\left[\delta_i = j\mid\ldots\right] \propto \begin{cases}
        n_j^{-i}\mathcal{PG}\left(\bm{y}_i\mid\bm{\alpha}_j,\bm{\beta}_j\right)\hspace{0.5cm} &\text{~for~}j \in \lbrace 1,\ldots,J^{-i}\rbrace\\
        \frac{\eta}{m}\mathcal{PG}\left(\bm{y}_i\mid\bm{\alpha}_j,\bm{\beta}_j\right)\hspace{0.5cm} &\text{~for~}j \in \lbrace J^{-i} + 1,\ldots, J^{-i} + m\rbrace.
        \end{cases}
\end{equation}
Of course, if a candidate cluster is selected, then $\delta_j = J + 1$, and $J = J^{-i} + 1$.

A key feature of the the projected Gamma distribution is its computational properties.  With respect to the distributional parameters $\bm{\alpha}_i,\bm{\beta}_i$, the projected Gamma density is proportional to
\begin{equation*}
    \mathcal{PG}(\bm{y}\mid\bm{\alpha}_i,\bm{\beta}_i) \propto \int\prod_{\ell = 1}^d\mathcal{G}\left(ry_{i\ell}\mid\alpha_{i\ell},\beta_{i\ell}\right)r^{d-1}dr,
\end{equation*}
which implies that, by reintroducing the latent radial component $r_i$ via sampling from its full conditional,
\begin{equation}
    r_i\mid\bm{\alpha}_i,\bm{\beta}_i \sim \mathcal{G}\left(\sum_{\ell = 1}^d\alpha_{i\ell}, \sum_{\ell = 1}^d\beta_{\ell} y_{i\ell}\right),
\end{equation}
the likelihood for $\bm{\alpha},\bm{\beta}$ becomes separable by dimension.  We stress this point: \emph{given} $r$, the distributional parameters $(\alpha_j,\beta_j)_{\ell}$ are independent in the likelihood.

The full conditional for $\bm{\alpha}_j,\bm{\beta}_j$ is then proportional to
\begin{equation}
    f(\bm{\alpha}_j,\bm{\beta}_j\mid \bm{Y},\bm{r},\bm{\delta},\ldots) \propto \prod_{i:\gamma_i = j}\prod_{\ell = 1}^d\mathcal{G}\left(r_iy_{i\ell}\mid\alpha_{j\ell},\beta_{j\ell}\right) \times dG_0(\alpha_{\ell},\beta_{\ell})
\end{equation}
The first form of the centering distribution we explore is a product of independent Gammas.
\begin{equation}
    G_0(\bm{\alpha}_j,\bm{\beta}_j\mid \bm{\xi},\bm{\tau},\bm{\zeta},\bm{\sigma}) = \prod_{\ell = 1}^d\mathcal{G}(\alpha_{j\ell}\mid \xi_{\ell},\tau_{\ell})\times\prod_{\ell = 2}^d\mathcal{G}(\beta_{j\ell}\mid\zeta_{\ell},\sigma_{\ell})
\end{equation}
This model is completed with Gamma priors on $\xi_{\ell}$, $\tau_{\ell}$, $\zeta_{\ell}$, $\sigma_{\ell}$.  Optionally, placing a Gamma prior on $\eta$ allows updating via the procedure outlined in \cite{escobar1995}.  In further analysis, we refer to this model as the \emph{projected Gamma--Gamma} (PG--G) model.  An advantage to using a Gamma centering distribution for the rate parameter is, as a conjugate prior, the rate parameters can easily be integrated out for inference on $\bm{\alpha}_j$.  After doing so, the full conditional for $\alpha_{j\ell}$ takes the form
\begin{equation}
    \pi(\alpha_{j\ell}\mid \bm{r},\bm{Y},\bm{\gamma},\xi_\ell,\tau_\ell,\zeta_\ell,\sigma_\ell) \propto \frac{\left(\prod_{i:\gamma_i = j}r_iy_{i\ell}\right)^{\alpha_{j\ell} - 1}\alpha_{j\ell}^{\xi_\ell - 1}e^{-\tau_\ell \alpha_{j\ell}}\times\Gamma\left(n_j\alpha_{j\ell} + \zeta_{\ell}\right)}{\Gamma^{n_j}(\alpha_{j\ell})\left(\sum_{i:\gamma_i = j}r_iy_{i\ell} + \sigma_{\ell}\right)^{n_j\alpha_{j\ell} + \zeta_{\ell}}}
\end{equation}
for $\ell = 2,\ldots,d$.  For $\ell = 1$, as $\beta_{1} := 1$, the full conditional takes the form
\begin{equation}
    \label{eqn:alpha1update}
    \pi(\alpha_{j1}\mid\bm{r},\bm{Y},\bm{\gamma},\xi_1,\tau_1) \propto \frac{\left(\prod_{i:\gamma_i = j}r_iy_{i1}\right)^{\alpha_{j1} - 1}\alpha_{j1}^{\zeta_1 - 1}e^{-\tau_1\alpha_{j1}}}{\Gamma^{n_j}(\alpha_{j1})}.
\end{equation}
Updating these parameters can be done via a simple Metropolis-within-Gibbs step.  In our analysis, we first transform $\alpha_{j\ell}$ to log scale, and use a normal proposal density.
The full conditional for $\beta$ is then
\begin{equation}
    \beta_{j\ell}\mid\bm{r},\bm{Y},\alpha,\zeta_{\ell},\sigma_{\ell} \sim \mathcal{G}\left(\beta_{j\ell}\mid n_j\alpha_{j\ell} + \zeta_\ell, \sum_{i:\gamma_i = j}r_iy_{i\ell} + \sigma_{\ell}\right),
\end{equation}
for $\ell = 2,\ldots, d$.  Updating $\beta_{j\ell}$ is done via a Gibbs step.  The hyper-parameters $\xi_{\ell},\tau_{\ell},\zeta_{\ell},\sigma_{\ell}$ follow similar Gamma-Gamma update relationships.  We also explore a restricted form of this model, where $\beta_{\ell} = 1$ for all $\ell$.  Under this model, we use the full conditional in Equation~\ref{eqn:alpha1update} for all $\ell$, and omit inference on $\bm{\zeta},\bm{\sigma}$.  We refer to this model as the \emph{projected restricted Gamma--Gamma} (PRG--G) model.

The second form of centering distribution we explore is a multivariate log-normal distribution on the shape parameters $\bm{\alpha}_j$, with independent Gamma $\beta_{j\ell}$ rate parameters.  
\begin{equation}
    G_0\left(\bm{\alpha}_j,\bm{\beta}_j\mid\bm{\mu},\bm{\Sigma},\zeta,\sigma\right) = \mathcal{LN}\left(\bm{\alpha}_j\mid\bm{\mu},\bm{\Sigma}\right)\times\prod_{\ell = 2}^d\mathcal{G}\left(\beta_{j\ell}\mid\zeta_{\ell},\sigma_{\ell}\right)
\end{equation}
This model is completed with a normal prior on $\bm{\mu}$, an inverse Wishart prior on $\bm{\Sigma}$, and Gamma priors on $\zeta_{\ell}$ and $\sigma_{\ell}$, and $\eta$.  This model is denoted as the \emph{projected Gamma--log-normal} (PG--LN) model.  We also explore a restricted Gamma form of this model as above, where $\beta_{\ell} = 1$ for all $\ell$.  This is denoted as the \emph{projected restricted Gamma--log-normal} (PRG--LN) model.  in the interest of brevity, we omit the inferential forms for this model, sufficing to say that updating $\bm{\alpha}_j$ is accomplished with a joint Metropolis-within-Gibbs step; the normal prior for $\mu$ is conjugate for the log-normal $\log\bm{\alpha}_j$, and updated via a Gibbs step; and an inverse Wishart prior for $\Sigma$ is again conjugate to the log-normal $\bm{\alpha}_j$, meaning it too can be updated via a Gibbs step.

\subsubsection{Choice of Norm\label{subsubsec:norm}}
In the extreme value case, we have defined our task on ${\mathbb S}_{\infty}^{d-1}$, the positive orthant
  of the hypersphere defined under the $\mathcal{L}_{\infty}$ norm.  We previously described the
  difficulty in establishing a density directly on this geometry.  However, we can build the projected
  gamma distribution on ${\mathbb S}_p^{d-1}$ for a large $p$.  We can then project samples from the
  posterior predictive distribution for a model built on ${\mathbb S}_p^{d-1}$ onto
  ${\mathbb S}_{\infty}^{d-1}$. As $p$ increases, the closer the manifold will be to
  ${\mathbb S}_{\infty}^{d-1}$, and thus the closer the projected distribution will be to an idealized
  projected Gamma built on ${\mathbb S}_{\infty}^{d-1}$.  In testing, we tried developing models on
  $S_p^{d-1}$ for varying levels of $p$, and found that higher $p$ produced higher performing models.
  Additionally, the difference in model performance became more apparent as dimensionality increased.
  We present results here for models developed on ${\mathbb S}_{10}^{d-1}$.
